Reanalysis Step 2: Multiple Imputation

### 2025-08-30 reanalysis-2-multiple_imputation.qmd
attempt 2l.pan

df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number()
  ) |>
  # Join the person-level DR severity
  left_join(df_person_dr, by = "e_mrn_deidentified") |>
  dplyr::select(row_id, dplyr::all_of(imputation_vars), person_dr_max) |>
  mutate(
    # Convert person_dr_max to ordered factor for polr imputation method
    person_dr_max = factor(person_dr_max, levels = c(0, 1, 2), 
                          labels = c("No_DR", "NPDR", "PDR"),
                          ordered = TRUE),
    # Convert original person_dr to factor (keep for comparison but won't impute)
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    cohort_id = factor(cohort_id),  # Ensure cohort_id is a factor
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat),

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for IDs)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor; see below.
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor; see below.

@@

# Don't use IDs as predictors AND don't impute them
# EXCEPT: e_mrn_deidentified is set to -2 as cluster indicator for 2l methods
pred_matrix[, "e_mrn_deidentified"] <- -2  # Cluster indicator for 2l.pan
pred_matrix["e_mrn_deidentified", ] <- 0  # Don't predict e_mrn itself
pred_matrix[, "row_id"] <- 0
pred_matrix["row_id", ] <- 0

>>>>

df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number(),
    # Create numeric cluster ID for 2l.pan (required for multilevel imputation)
    cluster_id = as.numeric(factor(e_mrn_deidentified))
  ) |>
  # Join the person-level DR severity
  left_join(df_person_dr, by = "e_mrn_deidentified") |>
  dplyr::select(row_id, cluster_id, dplyr::all_of(imputation_vars), person_dr_max) |>
  mutate(
    # Convert person_dr_max to ordered factor for polr imputation method
    person_dr_max = factor(person_dr_max, levels = c(0, 1, 2), 
                          labels = c("No_DR", "NPDR", "PDR"),
                          ordered = TRUE),
    # Convert original person_dr to factor (keep for comparison but won't impute)
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    cohort_id = factor(cohort_id),  # Ensure cohort_id is a factor
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat),

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for IDs)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor; see below.
custom_methods["cluster_id"] <- ""  # Cluster ID for 2l methods - never impute
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor; see below.

@@

# Don't use IDs as predictors AND don't impute them
# EXCEPT: cluster_id is set to -2 as cluster indicator for 2l methods
pred_matrix[, "cluster_id"] <- -2  # Cluster indicator for 2l.pan
pred_matrix["cluster_id", ] <- 0  # Don't predict cluster_id itself
pred_matrix[, "e_mrn_deidentified"] <- 0  # Don't use patient ID as predictor
pred_matrix["e_mrn_deidentified", ] <- 0  # Don't predict patient ID
pred_matrix[, "row_id"] <- 0
pred_matrix["row_id", ] <- 0
